---
title: "Getting Started with 'stansum'"
output: 
  rmarkdown::html_vignette:
    toc: true
bibliography: refs.bib
vignette: >
  %\VignetteIndexEntry{Getting Started with 'stansum'}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(stansum)
```


Package **stansum** provides a collection of models implemented in Stan [@stan-manual] and prepared for R using **instantiate** [@r-instantiate]. What follows is a rudimentary rundown of how to use the package.

The list of implemented models can be inspected with `list_models()`. At this moment the following models are implemented:

```{r list-models, results="asis", echo=FALSE}
paste0("- `",list_models(), "`") |> 
  cat(sep = "\n")
```

The recommended way of using the models is via the provided R wrappers.


# Fitting models included in the package

Let's load an example dataset:

```{r load-data}
data("Fake_maltiel_RD", package="stansum")
str(Fake_maltiel_RD)
```

It is an ARD consisting of 1000 respondents and 3 sub-populations sampled from the Random Degree model of @maltiel_etal_2015.




## @maltiel_etal_2015 Barrier Effects model

```{r}
fit <- maltiel_bem_count(
  N = nrow(Fake_maltiel_RD),
  K = ncol(Fake_maltiel_RD),
  y = Fake_maltiel_RD,
  L = 0,
  m = c(100, 200, 300) / 10000,
  seed = 666
)
```





## A test Bernoulli model

All models have R wrappers. For example `test_model()` wraps a simple Bernoulli model borrowed from Stan documentation:

```{r test-model-fit}
y <- c(1,1,1,1, 0,0,0,0,0,0)
fit <- test_model(N = length(y), y = y, seed = 666, show_messages = FALSE)
```

The results are:

```{r test-model-see}
class(fit)
fit$summary()
```



# Loading models

Use `get_model()` to get a **cmdstanr**-style [@r-cmdstanr] object of R6 class "CmdStanModel", for example:

```{r get-model}
mod <- get_model("bernoulli")
class(mod)
mod
```




# References
